---
title: "Random Forests, LAs"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
# library(rgdal)
# library(rgeos)
# library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)
library(caret)
library(randomForest)
library(CAST)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd, lsoa11cd, msoa11cd, ladcd, ladnm) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, ladcd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>% # for n == 1 
  ungroup()

# Complete panel
## REPLACE WITH 0.1 INSTEAD OF OF 0 FOR THE GROWTH
df <- df %>% filter(!is.na(ladcd)) %>% 
  complete(ladcd, year) %>% 
  replace(is.na(.), 0)

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## Firms

```{r, eval=T}

# also saved locally in /data/raw/firm_data_backup.csv

firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         firms = OBS_VALUE)

df <- df %>% left_join(firms) %>% 
  mutate(n_by_firm = n/firms) %>% # N websites/firm
  #filter(!is.na(firms)) %>% 
  dplyr::select(-n, -firms) %>% 
  rename(n = n_by_firm)

# df %>% filter(is.na(n_by_firm))
# sapply(df %>% left_join(firms), function(x) sum(is.na(x)))
```

## Spatial data

```{r}
# get LA for the UK
path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
la <- st_read(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
la <- st_transform(la, 4326)

#la.f <- fortify(la, region = "LAD21CD")

# cities
cities.1 <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  slice_max(pop, n = 10)
cities.2 <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  filter(name == "Belfast")

cities.sf <- bind_rows(cities.1, cities.2) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326)
```

## Distances to cities and retail centres

First set up the destinations. Origins are all the LA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

```{r}
# LA centroids
# la.c <- gCentroid(la, byid=TRUE)@coords %>% 
#   as.data.frame() %>%
#   bind_cols(la@data) %>% 
#   dplyr::select(LAD21CD, x, y) %>% 
#   remove_rownames()

sf_use_s2(FALSE)
la.c <- st_centroid(la) %>% 
  st_coordinates() %>%
  as.data.frame() %>% 
  bind_cols(la %>% dplyr::select(LAD21CD))

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK",
"Belfast, UK")	

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
#retail <-readOGR(geo.path) 
retail <- st_read(geo.path)

#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
# retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
#   as.data.frame() %>% 
#   add_rownames(var = "id") #%>%
retail.major.cetres.help <- st_centroid(retail) %>% 
  st_transform(4326) %>%
  st_coordinates() %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# retail.major.cetres <- retail.major.cetres.help %>% 
#   st_as_sf(coords = c("x", "y"), crs = 27700) %>%
#   st_transform(4326) %>%
#   st_coordinates() %>%
#   as_tibble() %>% 
#   bind_cols(retail.major.cetres.help$id) %>% 
#   rename(id = '...3') %>% 
#   left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
retail.major.cetres <- retail.major.cetres.help %>% 
  # as_tibble() %>% 
  # bind_cols(retail.major.cetres.help$id) %>% 
  # #rename(id = '...3') %>% 
  left_join(retail %>% rownames_to_column(var = "id"), by = 'id')
```

Distance to cities

```{r}
dist <- distm(cbind(la.c$X, la.c$Y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()  

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names 

dist <- dist %>% bind_cols(la.c$LAD21CD) %>% 
  rename(ladcd = last_col()) %>% 
  relocate(ladcd)

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

dist <- transform(dist, dist = do.call(pmin, dist[-1]))



# NOT USED AS IT SCREWS UP THE REGRESSIONS
# Complete panel, N = 3871903, 17 year * 227759 OA
# df <- df %>% bind_rows(dist %>% select(oa11cd)) %>% 
#   mutate(year = ifelse(is.na(year),1000, year),
#          country = substring(oa11cd, 1, 1)) %>% 
#   filter(country != "L" & country != "M" & country != "N") %>% 
#   filter(!is.na(country)) %>% 
#   complete(year, oa11cd, fill = list(n = 0)) %>% 
#   filter(year!=1000,
#          !is.na(oa11cd)) %>% 
#   select(-country)



# tests
# test$country <- substring(test$oa11cd, 1, 1)
# unique(test$country)
# length(unique(test$oa11cd))

# calculate area
#la$SHAPE_Area <- area(la) #/ 1000000
la$area <- st_area(la) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
         density = n / area) %>% 
  relocate(density, .after = area)

df <- df %>% filter(!is.na(London))  # to drop 34 for L99999999 and M99999999

sapply(df, function(x) sum(is.na(x)))
```

Distance to retail centres

```{r}
dist.retail <- distm(cbind(la.c$X, la.c$Y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()  

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names 

dist.retail <- dist.retail %>% bind_cols(la.c$LAD21CD) %>% 
  rename(LAD21CD = last_col()) %>% 
  relocate(LAD21CD)

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% dplyr::select(LAD21CD, dist.retail)

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("ladcd" = "LAD21CD")) 

sapply(df, function(x) sum(is.na(x)))
```

## df for RF

```{r}
# London = Islington 
# sf_use_s2(FALSE)
# path.geo <- paste0(path, "/data/raw/Local_Authority_Districts_(December_2021)_UK_BUC.geojson")
# la.sf <- st_read(path.geo)

#help1 <- st_filter( la.sf, cities.sf) %>% as_tibble %>% dplyr::select(LAD21CD) %>% 
#  left_join(df, by = c("LAD21CD"="ladcd")) %>% dplyr::select(LAD21CD, year, n)



#************** DOUBLE CHECK THE ORDER OF CITY NAMES BELOW **************

# nearest city
help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Belfast",
                    "Edinburgh",
                    "Glasgow"))

# dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]
# help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
#   rename(nearest.city.LAD21CD = LAD21CD) %>% 
#   dplyr::select(ladcd, nearest.city.LAD21CD)

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

sf_use_s2(FALSE)
la.c <- st_centroid(la)
kn <- knearneigh(la.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd")) 

df <- df %>% left_join(df %>% dplyr::select(ladcd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.LAD21CD" = "ladcd",
                              "year" = "year")) %>% 
  dplyr::select(-c(LAD21NMW, OBJECTID, BNG_E, BNG_N, LONG, LAT, SHAPE_Length)) %>% 
  relocate(LAD21NM, .after = ladcd)

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(LAD21CD, RGN21CD)

df <- df %>% left_join(lookup.region, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(RGN21CD, .after=year)

# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("E12000006", "E12000007", "E12000008", "E12000009" )
df <- df %>% mutate(south = ifelse(RGN21CD %in% south, 1, 0))

# For growth:
#df <- df %>% mutate(growth = log(n/n.lag))                          
df <- df %>% mutate(growth = log((n+0.001)/(n.lag+0.01)))                       # CONSTANT FOR GROWTH
#df <- df %>% mutate(growth = n- n.lag)                                         # ABS. GROWTH

sapply(df, function(x) sum(is.na(x)))
```

## unregister_dopar function

This is helpful when the parallel loop collapses. 

```{r}
unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}
unregister_dopar()
```

## RF with CAST

```{r}
df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% left_join(df %>% 
                         ungroup() %>% 
                         filter(LAD21NM == "Islington") %>% 
                         dplyr::select(year, n) %>% 
                         rename(n.London = n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.London.lag = dplyr::lag(n.London)) %>% 
  mutate(n.nearest.city.lag = dplyr::lag(n.nearest.city)) %>% 
  ungroup()

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996)
sapply(df, function(x) sum(is.na(x)))
```

### train the model in all data

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
                                spacevar = "RGN21CD",
                                timevar = "year", 
                                k = 10)

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

# detectCores()
# cl <- makePSOCKcluster(14)
# registerDoParallel(cl)

start.time <- Sys.time()
model.all <- train(n ~                                                          # REPLACE n WITH growth
                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
                   data = df,# %>% filter(year > 1999),                         # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity",
                   num.threads = 15)

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

print(model.all)

# [reproducible results]
# Random Forest 
# 
# 5984 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 5082, 4354, 4650, 4746, 5130, 4928, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#   2     variance    0.03216513  0.7810704  0.02077849
#   2     extratrees  0.03220664  0.7752043  0.02087374
#   4     variance    0.03271488  0.7775723  0.02134707
#   4     extratrees  0.03229937  0.7760603  0.02102094
#   6     variance    0.03397678  0.7680207  0.02235307
#   6     extratrees  0.03248101  0.7767931  0.02132943
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule = variance
#  and min.node.size = 5.
 
varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

### FROM NOW ON, IT HAS NOT BEEN USED FOR SUBMISSION 2

### train the model in a loop for all but one regions

```{r}
rm(indices)

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
set.seed(831)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2

#for the last model
set.seed(832)
seeds[[11]]<-sample.int(1000, 1)

# detectCores()
cl <- makePSOCKcluster(14)
registerDoParallel(cl)

# train in every region
start.time <- Sys.time()

regions <- df %>% ungroup() %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  
  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i), #& year > 1999),   # FILTER BY YEAR FOR GROWTH 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)
  # CV
  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')


  model.all <- train(n ~                                                        # REPLACE n WITH growth
                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
}

stopCluster(cl)
end.time <- Sys.time() 
time.taken <- round(end.time - start.time,2)
time.taken # ca. 4 minutes 
```

### Resamples for plots

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

```{r}
help <- lapply(fit.model.all, function(x) varImp(x)) 
help <- lapply(help, function(x) x[[1]])
help <- do.call("rbind", help)

importance <- help %>% rownames_to_column() %>% 
  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
  rename(importance = Overall) %>% 
  relocate(importance, .after = year) %>% 
  dplyr::select(-rowname)

importance %>% ggplot() +
  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
  scale_color_discrete(name = "Features") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("Importance") + 
  ggtitle("Feature importance") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

```

### Test on one region

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
# train on all but one region, test on that region

df.predictions <- data.frame()
for (i in regions){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
                    df[df$RGN21CD==i,]) #& df$year > 1999,])                            # FILTER BY YEAR
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)
    
    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
    d <- d %>% dplyr::select(RGN21CD, ladcd, year, n, 1)                                # REPLACE n WITH growth
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD) 

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))       # REPLACE n WITH growth

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
#   geom_tile() +
#   coord_fixed() +
#   labs(y = "", x = "") + #y: test.region, x: train.region
#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
#   theme(#legend.position = "none",
#         axis.text.x = element_text(angle = 45, hjust=1),
#         panel.border = element_blank(), 
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_rect(fill = "white"))
# ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

## RF growth with CAST

### train the model in all data, growth

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # , 11) 9 after adding South

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')

detectCores()
cl <- makePSOCKcluster(14)
registerDoParallel(cl)

start.time <- Sys.time()
model.all <- train(growth ~                                                     # REPLACE n WITH growth
                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

print(model.all)

# [reproducible results]
# Random Forest 
# 
# 4862 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 4044, 3685, 4056, 4056, 4032, 3707, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE       Rsquared   MAE      
#   2     variance    0.2004565  0.6285613  0.1460545
#   2     extratrees  0.2001702  0.6342523  0.1480880
#   4     variance    0.2080782  0.6066952  0.1503476
#   4     extratrees  0.2065532  0.6220336  0.1527747
#   6     variance    0.2099617  0.5996347  0.1507628
#   6     extratrees  0.2084788  0.6180880  0.1544252
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's website density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of website density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's website density, t-1", rowname)))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA_growth.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

### train the model in a loop for all but one regions, growth

```{r}
rm(indices)

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
set.seed(831)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2

#for the last model
set.seed(832)
seeds[[11]]<-sample.int(1000, 1)

# detectCores()
cl <- makePSOCKcluster(14)
registerDoParallel(cl)

# train in every region
start.time <- Sys.time()

regions <- df %>% ungroup() %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  
  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN21CD!=i & year > 1999),      # FILTER BY YEAR FOR GROWTH 
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)
  # CV
  tc <- trainControl(method = "cv", #MOVED HERE< DOUBLE CHECK
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   index = indices$index,
                   savePredictions = 'final')


  model.all <- train(growth ~                                                   # REPLACE n WITH growth
                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN21CD!=i & year > 1999),              # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
}

stopCluster(cl)
end.time <- Sys.time() 
time.taken <- round(end.time - start.time,2)
time.taken # ca. 4 minutes 
```

### Resamples for plots, growth

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance, growth

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

```{r}
help <- lapply(fit.model.all, function(x) varImp(x)) 
help <- lapply(help, function(x) x[[1]])
help <- do.call("rbind", help)

importance <- help %>% rownames_to_column() %>% 
  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
  rename(importance = Overall) %>% 
  relocate(importance, .after = year) %>% 
  dplyr::select(-rowname)

importance %>% ggplot() +
  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
  scale_color_discrete(name = "Features") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("Importance") + 
  ggtitle("Feature importance") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

```

### Test on one region, growth

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
# train on all but one region, test on that region

df.predictions <- data.frame()
for (i in regions){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], 
                    df[df$RGN21CD==i & df$year > 1999,])                                # FILTER BY YEAR
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)
    
    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i & year > 1999))                         # FILTER BY YEAR
    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", tested.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$RGN21CD) 

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$growth))   # REPLACE n WITH growth

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

path.out <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
#   geom_tile() +
#   coord_fixed() +
#   labs(y = "", x = "") + #y: test.region, x: train.region
#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
#   theme(#legend.position = "none",
#         axis.text.x = element_text(angle = 45, hjust=1),
#         panel.border = element_blank(), 
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_rect(fill = "white"))
# ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

## Granger

```{r}
library(plm)
p.df <- pdata.frame(df, index= c("ladcd","year"))


test <- pgrangertest(n ~ n.l.slag, data = p.df, order = 1L) # *** reject the null that x does NOT cause y
# n ~ Lags(n, 1) + Lags(n.l.slag, 1), or
# n ~ Lags(n, 1) + Lags(n.slag, 2)
# n ~ Lags(n, 1)
test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)


test <- pgrangertest(n ~ n.slag, data = p.df, order = 1L)
# n ~ Lags(n, 1:1) + Lags(n.slag, 1:1)
# n ~ Lags(n, 1:1)
test$indgranger %>% as_tibble() %>% arrange(`p-value`) #filter(`p-value` < 0.01)


test <- pgrangertest(n.slag ~ n, data = p.df, order = 1L) # ** reject the null that x does NOT cause y
test$indgranger %>% as_tibble() %>% arrange(`p-value`)


p.df$n.London_dist = p.df$n.London/p.df$London
p.df$n.nearest.city_dist = p.df$n.nearest.city/p.df$dist

pgrangertest(n ~ n.London_dist, data = p.df, order = 1L)
pgrangertest(n ~ n.nearest.city_dist, data = p.df, order = 1L)

cor(p.df$n.London_dist, dplyr::lag(p.df$n.London_dist), use = "complete.obs")
```


