---
title: "Random Forests", OAs
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(rgeos)
library(sf)
library(spdep)
#library(tmap)
#library(maptools)
library(knitr)
library(REAT)
library(tidygeocoder)
library(geosphere)
library(broom)
#library(foreach)
library(doParallel)
library(raster)
library(plm)
#library(lmtest)
library(caret)
library(randomForest)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

## Load data

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  dplyr::select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd
```

The internet archive $1996$-$2010$ data is saved on /hdd.
The internet archive $2011$-$2012$ data is saved on ~/projects/web.diffusion/data/temp.

```{r}
n = 1 #number of unique postcodes. 
m = 12

data.folder <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 

data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df1112 <- read_csv(data.path2011_2012)

df <- bind_rows(df9610, df1112) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  #summarise(n = n()) %>%            # for multiple pc  
  summarise(n = sum(V1.domain)) %>%  # for n == 1 
  ungroup()

# Partially complete panel as not all OAs have a website in at least ine year
df <- df %>% filter(!is.na(oa11cd)) %>% 
  complete(oa11cd, year) %>% 
  replace(is.na(.), 0)

# tests for Scotland
# df %>% filter(substr(oa11cd, 1,1) =="S",
#               year==2010) %>% summarise(mean(n, na.rm = T))
# 
# df %>% filter(oa11cd=="S00090182")
# 
# test@data %>% filter(substr(id, 1,1) =="S",
#                      n!=0)
# 
# df %>% filter(substr(oa11cd, 1,1) =="W",
#               year==2003)
```

## spatial data

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
colnames(oa.ew@data) <- c('id', 'area')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

# # get OA for Scotland
# path.geo.sc <- paste0(path, "/data/raw/SG_DataZoneBdry_2011")
# oa.sc <- readOGR(dsn=path.geo.sc, layer = "SG_DataZone_Bdry_2011")
# # source: https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011
# 
# # spatial transformations
# oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))
# 
# # Scotland (follows same steps as EW, see notes above)
# oa.sc <- oa.sc[, 'DataZone']
# colnames(oa.sc@data) <- c('id')
# 
# # reassign the polygon IDs
# oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))
# 
# # check the CRS has changed correctly, and the data slot has shrink to only the ID
# summary(oa.sc)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
colnames(oa.sc@data) <- c('id', 'area')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- maptools::spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)

# oa.gb$geometry <- oa.gb$geometry %>%
#   s2::s2_rebuild() %>%
#   sf::st_as_sfc()

# get GB
path.gb <- "/hdd/internet_archive/archive/gis/Countries_December_2014_Full_Clipped_Boundaries_in_Great_Britain.shp"
gb <- readOGR(path.gb)
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
```

## Full panel

```{r}
oa.sf <- st_as_sf(oa.gb)

oa.full <- replicate(17, oa.sf %>% as_tibble(), simplify = FALSE) %>%
  bind_rows(.id = 'id_') %>% 
  rename(year = id_) %>% 
  mutate(year = as.numeric(year) + 1995) %>% 
  dplyr::select(id, year)

df <- oa.full %>% left_join(df, by = c("id" = "oa11cd", "year" = "year")) %>% 
  replace(is.na(.), 0)

sapply(df, function(x) sum(is.na(x)))
rm(oa.full)
```

## Distances to cities and retail centres

First set up the destinations. Origins are all the OA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

```{r}
# oa centroids
# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())
oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK")	

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
retail <-readOGR(geo.path) 
#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id") #%>%

retail.major.cetres <- retail.major.cetres.help %>% 
  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble() %>% 
  bind_cols(retail.major.cetres.help$id) %>% 
  rename(id = '...3') %>% 
  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
```

Distance to cities

```{r}
dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()  

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names 

dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

dist <- transform(dist, dist = do.call(pmin, dist[-1]))

# NOT USED AS IT SCREWS UP THE REGRESSIONS
# Complete panel, N = 3871903, 17 year * 227759 OA
# df <- df %>% bind_rows(dist %>% dplyr::select(oa11cd)) %>%
#   mutate(year = ifelse(is.na(year),1000, year),
#          country = substring(oa11cd, 1, 1)) %>%
#   filter(country != "L" & country != "M" & country != "N") %>%
#   filter(!is.na(country)) %>%
#   complete(year, oa11cd, fill = list(n = 0)) %>%
#   filter(year!=1000,
#          !is.na(oa11cd)) %>%
#   dplyr::select(-country)

# tests
# test$country <- substring(test$oa11cd, 1, 1)
# unique(test$country)
# length(unique(test$oa11cd))

# calculate area
oa.gb$area <- area(oa.gb) #/ 1000000

# Join with distance and area
df <- df %>% left_join(dist, by = c("id" = "oa11cd")) %>% 
  rename(oa11cd=id) %>% 
  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)

sapply(df, function(x) sum(is.na(x)))
```

Distance to retail centres

```{r}
dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()  

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names 

dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% dplyr::select(oa11cd, dist.retail)

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 

sapply(df, function(x) sum(is.na(x)))
```

## df for RF

```{r}
## THIS DROPS L999999... AND NI, WHICH NEEDS TO BE FIXED
#df <- df %>% filter(!is.na(area)) 

# London = Islington 
cities.sf <- st_as_sf(cities, coords = c("long", "lat"))
cities.sf <- st_set_crs(cities.sf, 4326)

# help1 <- st_filter(oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
#   left_join(df, by = c("id"="oa11cd")) %>% dplyr::select(id, year, n)

sf_use_s2(FALSE)
help1 <- st_filter( oa.sf, cities.sf) %>% as_tibble %>% dplyr::select(id) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))
  
# too slow
#dist$dist.city.name <- names(dist)[apply(dist[,2:11], MARGIN = 1, FUN = which.min)]

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])

help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.oa11cd = id) %>% 
  dplyr::select(oa11cd, nearest.city.oa11cd)

#####save this#####
sf_use_s2(FALSE)
oa.c <- st_centroid(oa.sf)
kn <- knearneigh(oa.c, k = 5)
knn <- knn2nb(kn, row.names = NULL, sym = FALSE)
knn.l <- nb2listw(knn)
##################

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(oa11cd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("oa11cd" = "oa11cd")) 

df <- df %>% left_join(df %>% dplyr::select(oa11cd,year, n) %>% 
                         rename(n.nearest.city = n), 
                       by = c("nearest.city.oa11cd" = "oa11cd",
                              "year" = "year")) 

# Lookup: OA to regions
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
  mutate(help = substring(oa11cd, 1, 1),
         RGN11CD = ifelse(help=="E", RGN11CD,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", NA))),
         RGN11NM = ifelse(help=="E", RGN11NM,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", NA)))) %>% 
  dplyr::select(-help)

# df %>% ungroup() %>% distinct(RGN11NM)
# sapply(df, function(x) sum(is.na(x)))


# North/South
#lookup.region %>% distinct(RGN21NM, .keep_all = T)
south <- c("London", "South West", "South West", "East of England")
df <- df %>% mutate(south = ifelse(RGN11NM %in% south, 1, 0))

# For growth:
#df <- df %>% mutate(growth = log(n/n.lag))

sapply(df, function(x) sum(is.na(x)))

# out.path <- paste0(path, "/data/temp/oa_rf.RData")
# save.image(file = out.path)
# load(out.path)

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol", 
#                         "Manchester", "Sheffield", "Leeds", "Edinburgh", 
#                         "Leicester", "nearest.city.oa11cd")) %>% 
#   write_csv(file = out.path)
```

## RF

### train on all but one regions

```{r}
# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)

unregister_dopar <- function() {
  env <- foreach:::.foreachGlobals
  rm(list=ls(name=env), pos=env)
}
unregister_dopar()

# CV
tc <- trainControl(method = "cv",
                   number = 10, #10,
                   allowParallel = T,
                   savePredictions = 'final')

# London region: E12000007

# detectCores()
# cl <- makePSOCKcluster(14)
# registerDoParallel(cl)

# train
start.time <- Sys.time()

set.seed(71)
m.test <- train(n ~ density + area + dist + dist.retail + London +
                  n.slag + year + n.nearest.city, #n.l.slag
                data = df %>% filter(RGN11CD=="E12000008"),
                trControl = tc,
                method = "ranger", #"rf", 
                importance = "impurity",  #TRUE,
                na.action = na.omit)

#stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #1h

summary(m.test)
plot(m.test)
m.test
print(m.test)

varimp_mars <- varImp(m.test)
plot(varimp_mars, main="Variable Importance")
```

### test on the hold out regions
 
Wales region is missing!!!

```{r}
pred <- predict(m.test, df %>% ungroup() %>% filter(RGN11CD=="E12000006"), na.action = na.pass)
length(pred)

df %>% ungroup() %>% filter(RGN11CD=="E12000006") %>% filter(if_any(everything(), is.na))

pred <- cbind(pred %>% as_tibble(), df %>% ungroup() %>% filter(RGN11CD=="E12000006"))
postResample(pred = pred$value, obs = pred$n) ## 0.99
```

## RF rolling

### train the model in a loop

```{r}
# Train on t & t + 1

# Make the parallelisation reproducible using https://stackoverflow.com/questions/13403427/fully-reproducible-parallel-models-using-caret. 
# See also https://stackoverflow.com/questions/27944558/set-seed-parallel-random-forest-in-caret-for-reproducible-result

#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

detectCores()
cl <- makePSOCKcluster(6)#14
registerDoParallel(cl)

# run the model for every 2 years
start.time <- Sys.time()
years <- 2001:2012 #1997

for (t in years){
  
  tplus1 <- t + 1
  set.seed(71)
  print(t)
  
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.l.slag + n.nearest.city + year, #n.lag +
                     data = df %>% filter(year==t | year==tplus1), #%>% 
                       # mutate(growth = ifelse(growth == -Inf | growth == Inf, NA, growth)) %>% 
                       # mutate(growth = ifelse(is.na(growth), 0, growth)), 
                     #%>% mutate(growth = ifelse(growth== -Inf | growth == Inf, NA, growth))
                     trControl = tc,
                     method = "ranger", 
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = 'impurity') #TRUE
  
  #assign(paste0("model.all",as.character(t + 1)), model.all)
  file <- paste0(path, "/data/temp/oa_rf/timeseries", as.character(t + 1), ".RData")
  save(model.all, file = file)
  rm(model.all)
}

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

# df %>% filter(growth == Inf) 
# log(n/n.lag)

# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
```

### Resamples for plots

```{r}
# create a list of the above model outputs

load(file = (paste0(path, "/data/temp/oa_rf/timeseries2001.RData")))

filenames <- list.files(paste0(path, "/data/temp/oa_rf"), pattern = "*.RData", full.names = T)

t <- 1997
for (i in filenames){
  load(file = i)
  t <- t + 1
  assign(paste0("timeseries",t), model.all)  
}

fit.model.all <- mget(ls(pattern = "^timeseries\\d"))
rm(list = ls(pattern = "^timeseries\\d"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

```{r}
help <- lapply(fit.model.all, function(x) varImp(x)) 
help <- lapply(help, function(x) x[[1]])
help <- do.call("rbind", help)

importance <- help %>% rownames_to_column() %>% 
  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
  rename(importance = Overall) %>% 
  relocate(importance, .after = year) %>% 
  dplyr::select(-rowname)

importance %>% ggplot() +
  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
  scale_color_discrete(name = "Features") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("Importance") + 
  ggtitle("Feature importance") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

```

### test

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# growth or n

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
## Test on t + 2

# It will create: 2001 (train on 2000-2001) - 2011 (train on 2010)
# I need: 2001 (train on 2000-2001) - 2009 (train on 2008-2009)

# use the model trainned in year t to predict io of year t + 1
years <- (1998:2011)
pred.by.year.all <- data.frame(matrix(NA,nrow=227759, ncol = 0))
for (t in years){
  t_plus_1 <- t + 1
  pred <- predict(fit.model.all[names(fit.model.all)==paste0("timeseries",t)], df[df$year==t_plus_1,])
  pred <- as.data.frame(pred)
  rownames(pred) <- c()
  pred.by.year.all <- cbind(pred.by.year.all, pred)
}

#change column names to match with total
names(pred.by.year.all) <- 1999:2012

# wide to long
pred.by.year.all <- gather(pred.by.year.all,key = year,value = predictions,1:14)

# drop 2000 and select variables form total
df.no1997_1 <- df %>% 
  filter(year!=1996 & year!=1997 & year!=1998) %>% 
  dplyr::select(oa11cd, year, n) # CHANGE n WITH growth

# column bind prediction and data  
pred.by.year.all <- cbind(pred.by.year.all, df.no1997_1 %>% arrange(year, oa11cd)) #DOUBLE CHECK ARRANGE

# pred.by.year to list by year
pred.by.year.all.list <- split(pred.by.year.all, pred.by.year.all$year)

# calculate metrics for every year
rf.year.all.metrics <- lapply(pred.by.year.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth
rf.year.all.metrics %>% as.data.frame() %>%
  round(2) %>%
  kable()

# there are 2 columns named year, so I drop one
pred.by.year.all$year <- NULL

ggplot(data = pred.by.year.all,aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ year,ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```

## RF spatial 

```{r}
df <- df %>% mutate(help = substr(ladcd, start = 1, stop = 2)) %>% 
  mutate(RGN21CD = ifelse(is.na(RGN21CD), help, RGN21CD)) %>% 
  ungroup() %>% 
  dplyr::select(-help)

df <- df %>% dplyr::select(-n.lag)

```

### train the model in a loop

```{r}
#length is = (n_repeats*nresampling)+1
seeds <- vector(mode = "list", length = 11)

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 9) # 9 after adding South

#for the last model
seeds[[11]]<-sample.int(1000, 1)

# CV
tc <- trainControl(method = "cv",
                   number = 10,
                   seeds = seeds,
                   allowParallel = T,
                   savePredictions = 'final')

detectCores()
cl <- makePSOCKcluster(14)
registerDoParallel(cl)

# run the model for every 2 years
start.time <- Sys.time()
regions <- df %>% distinct(RGN21CD)
regions <- as.vector(regions$RGN21CD)

for (i in regions){
  set.seed(71)
  model.all <- train(n ~ area + dist + dist.retail + London + south + #growth ~ 
                      n.slag + n.nearest.city + year, #n.lag +  + n.l.slag
                     data = df %>% filter(RGN21CD==i), trControl = tc,
                     method = "ranger",
                     na.action = na.omit,
                     #preProc = c("center", "scale"),
                     importance = 'impurity') #TRUE
  assign(paste0("model.all",as.character(i)), model.all)
}

stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken

# df %>% filter(growth == Inf) 
# log(n/n.lag)

# The loop creates: 1998 (train on 1997-1998) - 2013 (train on 2012)
# I need: 1998 (train on 1997-1998) - 2011 (train on 2010-2011)
```

### Resamples for plots

```{r}
# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# resamples
fit.model.all.res <- resamples(fit.model.all)
summary(fit.model.all.res)

# plot metrics

# TODO: fix the year labels

scales <- list(x=list(relation="free"), y=list(relation="free"))
bwplot(fit.model.all.res, scales=scales)

# TODO: fix the year labels
```

### Variable importance

```{r , include=TRUE, results= 'markup', message=FALSE}
lapply(names(fit.model.all), function(x) plot(varImp(fit.model.all[[x]]), main=as.character(x)))
```

```{r}
help <- lapply(fit.model.all, function(x) varImp(x)) 
help <- lapply(help, function(x) x[[1]])
help <- do.call("rbind", help)

importance <- help %>% rownames_to_column() %>% 
  mutate(feature = gsub("model.all[0-9]+.", "", rowname),
         year = gsub(".*?([0-9]+).*", "\\1", rowname)) %>% 
  rename(importance = Overall) %>% 
  relocate(importance, .after = year) %>% 
  dplyr::select(-rowname)

importance %>% ggplot() +
  geom_line(aes(x=year, y = importance, color = feature, group = feature), size = 1) +
  scale_color_discrete(name = "Features") +
  #scale_linetype_manual(name = "Website density regressions111") 
  ylab("Importance") + 
  ggtitle("Feature importance") +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        plot.title = element_text(hjust = 0.5))

```

### test

```{r , include=TRUE, results= 'markup', message=FALSE, fig.height=15, fig.width=10}

# growth or n

# for a reference point
df %>% group_by(year) %>%
  summarise(min = min(n), max=max(n),
            mean = mean(n), median = median(n)) %>%
  round(2) %>% kable()
```

```{r}
## Test on t + 2

# train on one region, test on all the rest

df.predictions <- data.frame()
for (i in regions){
  for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all",i)], df[df$RGN21CD==j,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", i, ".on.", j), pred)
    
    d <- paste0("region.predict.model.", i, ".on.", j)
    d <- cbind(get(d), df %>% filter(RGN21CD==j)) #%>% arrange(year, ladcd)
    d <- d[,c(5, 2, 4, 6, 1)]
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  }
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN21CD, ".", trained.on))

# split to list by region pair
pred.by.region.all.list <- split(df.predictions, df.predictions$test.train) #correct length = 132 = 12*11

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))  # CHANGE n WITH growth

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  add_row(RGN21CD = c("S1", "N0", "W0"), RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

rf.year.all.metrics %>% 
  as.data.frame() %>%
  add_rownames(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN21CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN21CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared)
  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### test plots

```{r}
#plots
ggplot(data = df.predictions, aes(x = n, y = predictions)) +
  geom_point(colour = "blue") +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  #geom_vline(xintercept = 23, colour = "green", linetype = "dashed") +
  facet_wrap(~ test.train, ncol = 2) +
  #coord_cartesian(xlim = c(0,50000),ylim = c(0,50000)) +
  ggtitle("Predicted vs. actual trade by year")
```












