---
title: "point.distance"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rprojroot)
library(rgdal)
library(spdep)
library(dbscan)

options(scipen=10000)

# This is the project path
path <- find_rstudio_root_file()
```

Postcode lookup.

[source](https://geoportal.statistics.gov.uk/datasets/postcode-to-output-area-to-lower-layer-super-output-area-to-middle-layer-super-output-area-to-local-authority-district-august-2021-lookup-in-the-uk/about)

```{r}
path.lookup <- paste0(path,"/data/raw/PCD_OA_LSOA_MSOA_LAD_AUG21_UK_LU.csv")
lookup <- read_csv(path.lookup) %>% 
  select(pcds, oa11cd) %>% 
  dplyr::rename(pc = pcds)
#glimpse(lookup)
# The problems refer to Welsh LAD names. Not a problem for the analysis.
#sapply(lookup, function(x) sum(is.na(x)))
# 10332 missing msoa11cd

# path.lookup <- "/hdd/internet_archive/archive/ONSPD_FEB_2017_UK.csv"
# lookup <- read_csv(path.lookup) %>% 
#   select(pcds, oseast1m, osnrth1m) %>% 
#   dplyr::rename(pc = pcds)

```

The internet archive data is saved on /hdd.

```{r}
n = 1 #number of unique postcodes 

data.folder9610 <- "/hdd/internet_archive/archive/data/"
data.path9610 <- paste0(data.folder9610, "domain_pc_year.csv")
#Created by domain.R, which uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

data.path1112 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

df9610 <- read_csv(data.path9610) 
df1112 <- read_csv(data.path1112)

df <- bind_rows(df9610, df1112) %>% 
  filter(V1.domain == n,
         year > 1995) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) 

rm(df9610, df1112)
```

## Boundary data

```{r}
# get GB
path.gb <- paste0(path, "/data/raw/Countries_(December_2020)_UK_BUC.geojson")
gb <- readOGR(path.gb) 
# spatial transformations
gb <- spTransform(gb, CRS("+init=epsg:4326"))
gb <- subset(gb, CTRY20NM != "Northern Ireland")
gb.df <- fortify(gb, region = "CTRY20NM")

# cities
cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- readOGR(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- spTransform(oa.ew, CRS("+init=epsg:4326"))

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew[, c('OA11CD', 'Shape__Area')]
colnames(oa.ew@data) <- c('id', 'area')

# reassign the polygon IDs
oa.ew <- spChFIDs(oa.ew, as.character(oa.ew$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.ew)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- readOGR(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- spTransform(oa.sc, CRS("+init=epsg:4326"))

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc[, c('code', 'SHAPE_1_Ar')]
colnames(oa.sc@data) <- c('id', 'area')

# reassign the polygon IDs
oa.sc <- spChFIDs(oa.sc, as.character(oa.sc$id))

# check the CRS has changed correctly, and the data slot has shrink to only the ID
summary(oa.sc)

# build oa for GB
oa.gb <- spRbind(oa.ew, oa.sc)
rm(oa.ew, oa.sc)
```

## Distances to cities and retail centres

First set up the destinations. Origins are all the OA.

Major cities from [source](https://www.citypopulation.de/en/uk/cities/).
Retail centres from [source](https://data.cdrc.ac.uk/dataset/retail-centre-boundaries)

```{r}
# oa centroids
# oa.gb.c <- (rgeos::gCentroid(oa.gb, byid=TRUE))@coords %>% as.data.frame()
# oa.gb.c$id <- row.names(oa.gb.c)  #as_tibble(rownames = NA) %>% 
#   mutate(id = row.names())
oa.gb.c <- gCentroid(oa.gb, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id")

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK")	

cities <- geo(city.names, no_query = F, method = "arcgis")

# retail centres
geo.path <- paste0(path, "/data/raw/Retail_Boundaries_UK.gpkg")
retail <-readOGR(geo.path) 
#retail.major.cetres <- subset(retail, retail$Classification == "Major Town Centre") #%>% 
retail.major.cetres.help <- gCentroid(retail, byid=TRUE)@coords %>% 
  as.data.frame() %>% 
  add_rownames(var = "id") #%>%

retail.major.cetres <- retail.major.cetres.help %>% 
  st_as_sf(coords = c("x", "y"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble() %>% 
  bind_cols(retail.major.cetres.help$id) %>% 
  rename(id = '...3') %>% 
  left_join(retail@data %>% mutate(id = rownames(retail@data)), by = 'id')
```

Distance to cities

```{r}
dist <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(cities$long, cities$lat), fun=distHaversine) 
dist <- round((dist/1000),2) %>% 
  as_tibble()  

city.names <- city.names %>% stringr::str_remove(", UK")
names(dist) <- city.names 

dist <- dist %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# dist$dist <- names(dist)[apply(dist[-1], MARGIN = 1, FUN = which.min)]
# dist$distMet <- apply(dist[,2:11], 1, min)

dist <- transform(dist, dist = do.call(pmin, dist[-1]))

# tests
# test$country <- substring(test$oa11cd, 1, 1)
# unique(test$country)
# length(unique(test$oa11cd))

# Join with distance and area
```

Distance to retail centres

```{r}
dist.retail <- distm(cbind(oa.gb.c$x, oa.gb.c$y), cbind(retail.major.cetres$X, retail.major.cetres$Y), fun=distHaversine) 
dist.retail <- round((dist.retail/1000),2) %>% 
  as_tibble()  

# retail.names <- retail.major.cetres$RC_Name
# names(dist.retail) <- retail.names 

dist.retail <- dist.retail %>% bind_cols(oa.gb.c$id) %>% 
  rename(oa11cd = last_col()) %>% 
  relocate(oa11cd)

# Minimum distance 
dist.retail <- transform(dist.retail, dist.retail = do.call(pmin, dist.retail[-1]))
dist.retail <- dist.retail %>% select(oa11cd, dist.retail)
```

Joins

**IMPORTANT** these are distants not to points but to the centroids of the OA they fall within.

```{r}

df <- df %>% left_join(dist, by = c("oa11cd" = "oa11cd")) %>% 
  left_join(oa.gb@data, by = c("oa11cd" = "id")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))

# Join with complete panel
df <- df %>% left_join(dist.retail, by = c("oa11cd" = "oa11cd")) 

sapply(df, function(x) sum(is.na(x)))


```

## distance to London plot

```{r}
df %>%
  #mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=London, color=year, fill=year)) +
    geom_histogram(alpha=0.6, binwidth = 5) +
    #scale_fill_viridis(discrete=F) +
    #scale_color_viridis(discrete=F) +
    #theme_ipsum() +
    
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("domains") +
    facet_wrap(~year, scales = "free_y")
```

## distance to cities plot

```{r}
df %>%
  #mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=dist, color=year, fill=year)) +
  geom_histogram(alpha=0.6, binwidth = 5) +
    #scale_fill_viridis(discrete=F) +
    #scale_color_viridis(discrete=F) +
    #theme_ipsum() +
    
  # geom_histogram(aes(y = stat(count) / sum(count)), alpha=0.6, binwidth = 5) +
  # scale_y_continuous(labels = scales::percent)+

  
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("domains") +
    facet_wrap(~year, scales = "free_y")
```

## distance to retail centres

```{r}
df %>%
  #mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=dist.retail, color=year, fill=year)) +
    geom_histogram(alpha=0.6, binwidth = 5) +
    #scale_fill_viridis(discrete=F) +
    #scale_color_viridis(discrete=F) +
    #theme_ipsum() +
    
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("domains") +
    facet_wrap(~year, scales = "free_y")
```